Repdata
==========================================

## Load data and libraries
```{r load, results = "hide"}
activity <- read.csv("activity.csv")
library(dplyr)
library(ggplot2)
```


## total number of steps taken per day

#### histogram of total number of steps per day
``` {r hist}
unique(activity$date)
totalsteps <- activity %>% group_by(date) %>% summarise(total = sum(steps, na.rm = TRUE))
totalsteps
ggplot(totalsteps, aes(total)) + geom_histogram(binwidth = 2500, fill = blues9) + xlab("Total number of steps per day") + ylab("count of total number of steps per day") + ggtitle("Histogram of steps per day")
```  

#### mean and median number of total number of steps taken per day
``` {r mean and median}
mean(totalsteps$total, na.rm = TRUE)
median(totalsteps$total, na.rm = TRUE)
```

## Average daily pattern
```{r tsplot}
avesteps <- activity %>% group_by(interval) %>% summarise(average = mean(steps, na.rm = TRUE))
avesteps$interval <- (avesteps$interval)*100
ggplot(avesteps, aes(x= interval, y = average)) + geom_line(color = "steelblue", size = 1, alpha = 0.8)  
avesteps[which.max(avesteps$average),] #maximum number of steps lies at interval 8.35
```
## Imputing missing values

##### number of NA values in activity dataset
```{r missing}
colSums(is.na(activity)) 
```

#### replace missing values
NA values will be replaced by the average number of steps in the 5 minute interval
Create a new column complete_steps with all NA values replaced
```{r na}
activity$Complete_steps <- ifelse(is.na(activity$steps), round(avesteps$average[match(activity$interval, avesteps$interval)],0), activity$steps)
#merge 2 columns complete_steps and steps in the activity dataset
Finalactivity <- data.frame(steps = activity$Complete_steps, interval = activity$interval, date = activity$date)
head(Finalactivity, n = 5) #first 5 rows of new activity data

```

#### Histogram of total number of steps 
``` {r final}
Finalactivity_total <- Finalactivity %>% group_by(date) %>% summarise(total = sum(steps, na.rm = TRUE))
ggplot(Finalactivity_total, aes(total)) + geom_histogram(binwidth = 2500, fill = blues9)

```

#### mean and median total number of steps per day
``` {r}
mean(Finalactivity_total$total)
median(Finalactivity_total$total)
# there is not much impact on the mean and median.
```


## Differences in activity patterns between weekdays and weekends

``` {r}
library(lubridate)
activity$date <- as.Date(activity$date)
activity$day <- weekdays(activity$date)
unique(activity$day)
activity$daytype <- ifelse(activity$day == "Saturday" | activity$day == "Sunday",'Weekend','Weekday')
unique(activity$daytype)

activity_grouped <- activity %>% group_by(daytype, interval) %>% summarise(ave = mean(steps, na.rm = TRUE))

ggplot(activity_grouped, aes(x = interval, y = ave)) + geom_line(color = "steelblue") + facet_grid(daytype ~.) + ylab("average number of steps")


```







